
# Set up

Load packages
```{r}
library(tidyverse)
library(sf)
library(tmap)
library(rnaturalearth)
```

Settings and universal variables
```{r}
tmap_mode(mode = "view")

#minimum distance between USGS deposit and nearest mine (to ensure that we are considering unexploited deposits)
unexploited_dist = 100000 #100km

#max distance for cut-off between mines/deposits and Indigenous territories
dist_threshold = 10000 #10km (from Owen et al., 2022)

crs = 6933

options(scipen=999)

sf_use_s2(FALSE)
```

# Import data

## List of countries
```{r}
countries_df = read.csv("Processed data/data_breakdown_by_country_apr30.csv")

countries_full = as.vector(countries_df%>%filter((landmark %in% c("Full Data", "Indicative areas") | in_garnett == "Yes") & maus_polygons > 0)%>%pull(country))

countries_recog = as.vector(countries_df%>%filter(landmark == "Full Data" & maus_polygons > 0)%>%pull(country))
```

## Mine data
```{r}
snp = st_transform(st_read("Processed data/snp_points_filtered.gpkg"), crs)%>%
  filter(geounit %in% countries_full)

maus = st_transform(st_read("Processed data/maus_polygons_filtered_joined.gpkg"), crs)%>%
  filter(geounit %in% countries_full)
```

## ECM deposits
```{r}
usgs_deposits = st_transform(st_read("Processed data/usgs_deposits_filtered.gpkg"), crs)%>%
  filter(geounit %in% countries_full)%>%
  mutate(dist_to_maus = apply(st_distance(., maus), 1, min),
         dist_to_snp = apply(st_distance(., snp), 1, min),
         dist_to_mine = pmin(dist_to_maus, dist_to_snp),
         id = row_number())
  
usgs_deposits%>%
  ggplot(aes(dist_to_mine/1000))+
  geom_histogram(fill = "firebrick4", binwidth = 25)+
  theme_bw()+
  labs("Distance to Nearest Mine (km)")

summary(usgs_deposits$dist_to_mine/1000)
```

## Indigenous territories

2 landmark datasets: 1) Countries with landmark full data with recognition status and documentation, and 2) Countries with partial data for Landmark (polygon boundaries of territories but no recognition status)

Polygon boundaries of territories (no recognition status available) from Garnett et al

For now, clipping to a buffer near known deposits or mines to make the datasets smaller. Will also try Julie's suggestion of going from multipart to singlepart features to see if that helps.
```{r}
buffer = st_transform(rbind(st_buffer(usgs_deposits, 25*1000)%>%select(geom), st_buffer(maus, 25*1000)%>%select(geom), st_buffer(snp, 25*1000)%>%select(geom)), crs)

landmark_simplified = st_transform(st_read(dsn = "C:/Users/benji/Desktop/Landmark/landmark_arcgis/landmark_arcgis.gdb", layer = "community_landmark_simplified", options = "METHOD=SKIP"), crs)%>%
  filter(Country %in% countries_full)%>%
  rename(geometry = Shape)%>%
  st_filter(., buffer, .pred = st_intersects)%>%
  mutate(recognition = ifelse(Form_Rec == "Acknowledged by govt" & Doc_Status == "Documented", 1, 0))
 
landmark2_simp = st_transform(st_read(dsn = "C:/Users/benji/Desktop/Landmark/landmark_arcgis/landmark_arcgis.gdb", layer = "indicative_landmark_simplified", options = "METHOD=SKIP"), crs)%>%
  filter(Country %in% countries_full)%>%
  st_filter(., buffer, .pred = st_intersects)%>%
  rename(geometry = Shape)

garnett = st_transform(st_read("Raw data/Garnett et al/IPL_IndigenousPeoplesLands_2017/IPL_2017.shp"), crs)%>%
  left_join(., ne_countries(scale = "medium")%>%st_drop_geometry()%>%select(geounit, sov_a3), by = c("Name_" = "sov_a3"))%>%
  filter(geounit %in% countries_full)%>%
  select(geometry)%>%
  st_filter(., buffer, .pred = st_intersects)%>%
  mutate(Identity = NA, Doc_Status = NA, Form_Rec = NA)
  
landmark_full = rbind(landmark_simplified%>%select(Identity:Doc_Status), landmark2_simp%>%select(Identity:Doc_Status))

landmark_garnett = rbind(landmark_full, garnett)
```
Split up landmark into recognised and non-recognised lands. (Which variable(s) to split by subject to our group)
```{r}
landmark_recog = landmark_simplified%>%
  filter(recognition == 1)
landmark_no_recog = landmark_simplified%>%
  filter(recognition == 0)

landmark_simplified%>%
  mutate(total_area = sum(Area_GIS), totalarea2 = sum(Area_Ofcl), totaln = length(!is.na(Area_GIS)))%>%
  group_by(recognition)%>%
  reframe(area = sum(Area_GIS)/total_area, area2 = sum(Area_Ofcl)/totalarea2, n = n()/totaln)%>%
  unique()

rm(landmark_simplified, landmark2_simp, garnett)
```

# Proximities

Binary based on whether mine is within a certain distance of a territory or not
```{r}
snp%>%
  mutate(any = as.integer(lengths(st_intersects(st_buffer(., 10*1000), landmark_garnett)) > 0),
         recog = as.integer(lengths(st_intersects(st_buffer(., 10*1000), landmark_recog)) > 0),
         unrecog = as.integer(lengths(st_intersects(st_buffer(., 10*1000), landmark_no_recog)) > 0),)%>%
  data.frame()
```





# Distances

Calculate the minimum distance from each Maus polygon and unexploited USGS deposit to the nearest Indigenous/communal territory
```{r}
snp = snp%>%
  mutate(min_dist_all = as.numeric(st_distance(., landmark_garnett[st_nearest_feature(., landmark_garnett), ], by_element = TRUE)))

maus = maus%>%
  mutate(min_dist_all = as.numeric(st_distance(., landmark_garnett[st_nearest_feature(., landmark_garnett), ], by_element = TRUE)))

usgs_deposits = usgs_deposits%>%
  mutate(min_dist_all = as.numeric(st_distance(., landmark_garnett[st_nearest_feature(., landmark_garnett), ], by_element = TRUE)))
```



Distance to recognised and unrecognised territories
```{r}
snp = snp%>%
  left_join(., snp%>%
              filter(geounit %in% countries_recog)%>%
              mutate(min_dist_recog = as.numeric(st_distance(., landmark_recog[st_nearest_feature(., landmark_recog), ], by_element = TRUE)),
                     min_dist_unrecog = as.numeric(st_distance(., landmark_no_recog[st_nearest_feature(., landmark_no_recog), ], by_element = TRUE)))%>%
              select(LocalID, min_dist_recog, min_dist_unrecog)%>%
              st_drop_geometry(),
            by = "LocalID")

maus = maus%>%
  left_join(., maus%>%
              filter(geounit %in% countries_recog)%>%
              mutate(min_dist_recog = as.numeric(st_distance(., landmark_recog[st_nearest_feature(., landmark_recog), ], by_element = TRUE)),
                     min_dist_unrecog = as.numeric(st_distance(., landmark_no_recog[st_nearest_feature(., landmark_no_recog), ], by_element = TRUE)))%>%
              select(id, min_dist_recog, min_dist_unrecog)%>%
              st_drop_geometry(),
            by = "id")

usgs_deposits = usgs_deposits%>%
  left_join(., usgs_deposits%>%
              filter(geounit %in% countries_recog)%>%
              mutate(min_dist_recog = as.numeric(st_distance(., landmark_recog[st_nearest_feature(., landmark_recog), ], by_element = TRUE)),
                     min_dist_unrecog = as.numeric(st_distance(., landmark_no_recog[st_nearest_feature(., landmark_no_recog), ], by_element = TRUE)))%>%
              select(id, min_dist_recog, min_dist_unrecog)%>%
              st_drop_geometry(),
            by = "id")
```


Now, we can statistically test the differences in distances across these two datasets. 
```{r}
usgs_filtered = usgs_deposits%>%filter(dist_to_mine > 100*1000)

#versions for both maus and s&p mine datasets
combined_snp = bind_rows(snp%>%st_drop_geometry()%>%select(min_dist_all, min_dist_recog, min_dist_unrecog)%>%mutate(source = "ECM Mines"), 
                         usgs_filtered%>%st_drop_geometry()%>%select(min_dist_all, min_dist_recog, min_dist_unrecog)%>%mutate(source = "Unexploited ECM Deposits"))

combined_maus = bind_rows(maus%>%st_drop_geometry()%>%select(min_dist_all, min_dist_recog, min_dist_unrecog)%>%mutate(source = "ECM Mines"), 
                         usgs_filtered%>%st_drop_geometry()%>%select(min_dist_all, min_dist_recog, min_dist_unrecog)%>%mutate(source = "Unexploited ECM Deposits"))
```

Wilcoxon test to evaluate if mean distance for mines is less than unexploited deposits
```{r}
wilcox.test(min_dist_all ~ source, data = combined_snp, alternative = "less")
wilcox.test(min_dist_all ~ source, data = combined_maus, alternative = "less")

wilcox.test(min_dist_recog ~ source, data = combined_snp, alternative = "less")
wilcox.test(min_dist_recog ~ source, data = combined_maus, alternative = "less")

wilcox.test(min_dist_unrecog ~ source, data = combined_snp, alternative = "less")
wilcox.test(min_dist_unrecog ~ source, data = combined_maus, alternative = "less")
```
We don't observe any significant differences, except for unrecognised lands (both S&P and Maus) and recognised lands (Maus)


```{r}
# Two-proportion z-test to evaluate if mines are more likely to intersect Indigenous territories
prop.test(c(sum(snp$min_dist_all == 0), sum(usgs_filtered$min_dist_all == 0)), c(length(snp$min_dist_all), length(usgs_filtered$min_dist_all)), alternative = "greater")
prop.test(c(sum(maus$min_dist_all == 0), sum(usgs_filtered$min_dist_all == 0)), c(length(maus$min_dist_all), length(usgs_filtered$min_dist_all)), alternative = "greater")

# For min_dist_unrecog
prop.test(c(sum(snp$min_dist_unrecog == 0, na.rm=T), sum(usgs_filtered$min_dist_unrecog == 0, na.rm=T)), 
          c(sum(!is.na(snp$min_dist_unrecog)), sum(!is.na(usgs_filtered$min_dist_unrecog))), 
          alternative="greater")

prop.test(c(sum(maus$min_dist_unrecog == 0, na.rm=T), sum(usgs_filtered$min_dist_unrecog == 0, na.rm=T)), 
          c(sum(!is.na(maus$min_dist_unrecog)), sum(!is.na(usgs_filtered$min_dist_unrecog))), 
          alternative="greater")

# For min_dist_recog
prop.test(c(sum(snp$min_dist_recog == 0, na.rm=T), sum(usgs_filtered$min_dist_recog == 0, na.rm=T)), 
          c(sum(!is.na(snp$min_dist_recog)), sum(!is.na(usgs_filtered$min_dist_recog))), 
          alternative="greater")

prop.test(c(sum(maus$min_dist_recog == 0, na.rm=T), sum(usgs_filtered$min_dist_recog == 0, na.rm=T)), 
          c(sum(!is.na(maus$min_dist_recog)), sum(!is.na(usgs_filtered$min_dist_recog))), 
          alternative="greater")
```
For Maus, mines more likely to intersect for all territories, unrecognised, and recognised territories. 
For S&P, mines more likely to intersect for unrecognised territories. 


```{r}
wilcox.test(combined_snp$min_dist_unrecog, combined_snp$min_dist_recog, alternative = "less")
wilcox.test(combined_maus$min_dist_unrecog, combined_maus$min_dist_recog, alternative = "less")

prop.test(c(sum(snp$min_dist_recog == 0, na.rm=T), sum(snp$min_dist_unrecog == 0, na.rm=T)), 
          c(sum(!is.na(snp$min_dist_recog)), sum(!is.na(snp$min_dist_unrecog))), 
          alternative="greater")
prop.test(c(sum(maus$min_dist_recog == 0, na.rm=T), sum(maus$min_dist_unrecog == 0, na.rm=T)), 
          c(sum(!is.na(maus$min_dist_recog)), sum(!is.na(maus$min_dist_unrecog))), 
          alternative="greater")
```
Mines (for both S&P and Maus) on average are have lower distances to unreocgnised territories than recognised, but are less likely to intersect unrecognised territories



```{r}
#Visualise distance distributions
ggplot(combined_df, aes(x = source, y = min_dist_all, fill = source)) +
  geom_violin(trim = T, alpha = 0.5) +
  theme_minimal() +
  labs(x = "", y = "Distance to Indigenous/Communal Territories (km)", fill = "") +
  scale_fill_brewer(palette = "Set2")

ggplot(combined_df, aes(x = min_dist_all/1000, fill = source)) +
  geom_density(alpha = 0.4) +
  theme_minimal() +
  labs(x = "Distance to Indigenous/Communal Territories (km)", y = "Frequency", fill = "") +
  scale_fill_brewer(palette = "Set2")

ggplot(combined_df, aes(x = source, y = min_dist_all/1000, fill = source)) +
  geom_boxplot(alpha = 0.5) +
  theme_minimal() +
  labs(x = "", y = "Distance to Indigenous/Communal Territories (km)", fill = "") +
  scale_fill_brewer(palette = "Set2")
```
```{r}
#Visualise distance distributions
ggplot(combined_df, aes(x = source, y = min_dist_unrecog, fill = source)) +
  geom_violin(trim = T, alpha = 0.5) +
  theme_minimal() +
  labs(x = "", y = "Distance to Indigenous/Communal Territories (km)", fill = "") +
  scale_fill_brewer(palette = "Set2")
```


