
# Set up

Load packages
```{r}
suppressPackageStartupMessages({
  suppressWarnings({
    library(sf)          # Spatial data
  library(terra)       # Raster processing  
  library(spsurvey)    # Spatially balanced sampling
  library(MatchIt)     # Matching
  library(cobalt)      # Balance checking
  library(tidyverse)       # Data wrangling
  library(exactextractr) # Zonal stats
  library(rnaturalearth) # Country boundaries
  library(tmap)        # Making maps
  library(spsurvey)    # Spatially balanced sampling
  library(sandwich)
  library(lmtest)
  library(rmapshaper) # Spatial processing
  library(fixest) # fixed effects regression
  })})
```

Settings and universal variables
```{r}
tmap_mode(mode = "view")
tmap_options(check.and.fix = TRUE)

mine_buffer = 2500

crs = 4326
crs_sampling = 9822 #needs to be a projected crs, so chose an equal area global one

options(scipen=999)

set.seed(1236)
```

# Import & process data



## Covariates

Road density
```{r}
road_density_raw = rast("Raw data/Covariates/Road access/GRIP4_density_total/grip4_total_dens_m_km2.asc")
road_density = log(road_density_raw + 0.001)

rm(road_density_raw)
```

Distance to water
```{r}
dist_to_water = mean(rast("Raw data/Covariates/Distance to water/distance_to_water_reprojected.tif"))
```

Nightlights
```{r}
nightlights = rast("Raw data/Covariates/Nightlights/VNL_npp_2024_global_vcmslcfg_v2_c202502261200.average_masked.dat.tif/VNL_npp_2024_global_vcmslcfg_v2_c202502261200.average_masked.dat.tif")
```

Protected areas
```{r, eval = F}
protected_areas_full = st_read("Raw data/Covariates/Protected areas/protected_areas_study_area_full.shp")%>%
  filter(st_is_valid(geometry))

protected_areas = rast("Raw data/Covariates/Protected areas/protected.tif")
protected_areas[is.na(protected_areas)] = 0
```

Land cover
```{r}
land_cover_path = "Raw data/Covariates/Land cover/MCD12C1.A2023001.061.2024251212901.hdf"
#gdal_subdatasets(land_cover_path)

land_cover = project(rast(land_cover_path, subds = "Majority_Land_Cover_Type_1"), "EPSG:4326")

forest_binary = as.numeric(land_cover %in% 1:5)  # Classes 1-5 are forest types (Evergreen Needleleaf to Mixed Forests)
cropland_binary = as.numeric(land_cover %in% c(12,14)) # Class 12 is cropland, 14 is cropland mosaic

water = as.numeric(land_cover == 0)
water[water == 0] = NA 
water_sf <- st_as_sf(as.polygons(water, dissolve = TRUE))

rm(land_cover)
rm(water)
```

USGS ECM deposits
```{r}
usgs_deposits = st_transform(st_read("Processed data/Mines/usgs_deposits_filtered.gpkg"), crs)
```

Population Density
```{r}
pop_dens = rast("Raw data/Covariates/Population density/gpw_v4_population_density_rev11_2020_30_sec_2020.tif")
```


## Dependent variable (mines)
```{r}
snp = st_transform(st_read("Processed data/Mines/snp_points_filtered.gpkg"), crs)

maus = st_transform(st_read("Raw data/Maus et al/Global_Mining_Polygons_v2.shp"), crs)%>%
  mutate(maus_id = row_number())
#maus_alone = st_transform(st_read("Processed data/Mines/maus_polygons_filtered_joined.gpkg"), crs)%>%st_make_valid()
```

The maus polygons don't have any covariates in the dataset. However, whenever possible, we still want to get the extra mines from this dataset that are not in the S&P dataset. We do this by filtering for polygons that are 1) within 2.5km of an ECM deposit point (to ensure it’s an ECM mine), and 2) at least 2.5km from the nearest S&P point (to avoid duplicates). 
```{r}
maus = maus%>%
  st_filter(., st_buffer(usgs_deposits, mine_buffer), .predicate = st_intersects)

maus_near_snp = maus%>%
  st_filter(., st_buffer(snp, mine_buffer), .predicate = st_intersects)

maus_filtered = maus%>%
  filter(!maus_id %in% maus_near_snp$maus_id)

st_write(maus_filtered%>%select(geometry), "Processed data/Mines/maus_filtered_deduplicated.gpkg", delete_dsn = T)
rm(maus, usgs_deposits, maus_near_snp)

maus_filtered = st_read("Processed data/Mines/maus_filtered_deduplicated.gpkg")
```

## Independent variable (Landmark)

Load in our landmark datasets of Indigenous territories
```{r, eval = F}
#landmark_study_area_recog = st_transform(st_read("Processed data/Landmark processed/landmark_recog_study_area.shp"), crs)

landmark_study_area_full = st_transform(st_read("Processed data/Landmark processed/landmark_and_garnett_full_study_area.shp"), crs)%>%filter(st_is_valid(geometry))%>%
  mutate(
    documented = case_when(
      Doc_Status == "Documented" ~ 1,
      Doc_Status == "Unknown" | is.na(Doc_Status) ~ NA,
      T ~ 0),
    recognised = case_when(
      Form_Rec == "Acknowledged by govt" ~ 1,
      Form_Rec == "Unknown" | is.na(Form_Rec) ~ NA,
      T ~ 0),
    doc_rec = ifelse(documented == 1 & recognised == 1, 1, 0),
    indigenous = 1)

table(landmark_study_area_full$Doc_Status, landmark_study_area_full$Form_Rec)
table(landmark_study_area_full$Category)

# landmark_study_area_full_no_garnett = st_transform(st_read("Processed data/Landmark processed/landmark_full_study_area_no_garnett.shp"), crs)%>%filter(st_is_valid(geometry))

#garnett = st_transform(st_read("Raw data/Garnett et al/garnett_singlepart.shp"), crs)
```


## Study area

Define study area (based on adm1). We only want administrative areas that have at least one mine and one Indigenous territory, for both all territories and territories for which we have recognition data. 

So we create some new binary columns recording whether each adm1 intersects with a mine, indigenous territories, etc.
```{r}
adm1s <- st_transform(ne_download(scale = 10, type = "states", category = "cultural", returnclass = "sf"), crs)%>%
  st_make_valid()%>%
  mutate(geonunit = ifelse(is.na(geonunit), admin, geonunit),
         gn_name = ifelse(is.na(gn_name), name, gn_name))%>%
  select(gn_name, geonunit)%>%
  mutate(
    adm1_id = as.factor(row_number()),
    maus = (st_intersects(geometry, st_centroid(maus_filtered)) %>% lengths() > 0) %>% as.integer(),
    snp  = (st_intersects(geometry, snp) %>% lengths() > 0) %>% as.integer(),
    garnett = 0, landmark_community = 0, landmark_indicative = 0)
```

Due to computational constraints and broken geometries that sf struggles to handle, we're going to export this layer to ArcGIS Pro and create the columns for intersections with Indigenous territories there. 
```{r, eval = F}
st_write(adm1s, "Processed data/Study area/Intermediate/Adm1s, export to arcgis/adm1s_arcgis_export.shp", delete_dsn = T)
```

After doing that processing in ArcGIS Pro, we can now load our processed dataset back and filter for only adm1s that both contain ECM mines and intersect with Indigenous territories
```{r}
adm1s_arcgis = st_read("Processed data/Study area/Intermediate/Adm1s, export to arcgis/adm1s_arcgis_export.shp")%>%
  select(adm1_id, garnett, lndmrk_c, lndmrk_n)%>%
  rename(landmark_indicative = lndmrk_n, landmark_community = lndmrk_c)%>%
  st_drop_geometry()

adm1s_full = adm1s%>%select(-c(garnett, landmark_community, landmark_indicative))%>%
  left_join(., adm1s_arcgis, by = "adm1_id")%>%
  mutate(any_mine = ifelse(maus > 0 | snp > 0, 1, 0), 
         any_landmark = ifelse(landmark_indicative + landmark_community > 0, 1, 0),
         any_indig = ifelse(garnett + any_landmark > 0, 1, 0))

st_write(adm1s_full, "Processed data/Study area/adm1s_full.gpkg", delete_dsn = T)
rm(adm1s, adm1s_arcgis)
```

```{r}
adm1s_full = st_read("Processed data/Study area/adm1s_full.gpkg")

adm1s_filtered = adm1s_full%>%
  filter(any_mine == 1 & any_indig == 1)%>%
  filter(gn_name != "Chukotskiy Avtonomnyy Okrug")

adm1s_no_garnett = adm1s_filtered%>%
  filter(any_landmark == 1)

rm(adm1s_full)
```

Visualisation of study adm1s
```{r, eval = F}
tm_shape(adm1s_filtered)+
  tm_polygons(col = "coral", alpha = 0.5)

nrow(adm1s_filtered)
length(unique(adm1s_filtered$geonunit))
```
446 adm1s across 79 countries

Visualisation of study adm1s with recognition data
```{r}
tm_shape(adm1s_filtered%>%filter(landmark_community == 1))+
  tm_polygons(col = "#503570", alpha = 0.5)

nrow(adm1s_filtered%>%filter(landmark_community == 1))
length(unique(adm1s_filtered%>%filter(landmark_community == 1)%>%pull(geonunit)))
```


## Mines in study area

Let's filter our mine datasets (both Maus polygons and S&P points) to ensure that we're only considering mines that intersect the study area. 

For Maus, we'll also consolidate any duplicates based on the 'Local ID' column as well, and conduct a spatial join to bring in the adm1 and country data. 
```{r}
maus_study_area = maus_filtered%>%
  st_filter(., adm1s_filtered, .predicate = st_intersects)%>%
  st_join(., adm1s_filtered%>%select(adm1_id, gn_name, geonunit), largest = T)
  
st_write(maus_study_area, "Processed data/Mines/maus_study_area.gpkg", delete_dsn = T)
rm(maus_filtered)
maus_study_area = st_read("Processed data/Mines/maus_study_area.gpkg")
```

Now, for the S&P dataset, we also filter for points in the study area, and get rid of any duplicates, as defined by any mining points that are both 1) within or right next to (within 1km) a Maus mining polygon, AND 2) share the same primary commodity as the Maus polygon they're in/next to.
```{r}
snp_study_area = snp%>%
  rename(primary_commodity = Primary.Co)%>%
  select(primary_commodity)%>%
  #mutate(snp_id = row_number())%>%
  st_filter(., adm1s_filtered, .predicate = st_intersects)%>%
  st_join(., adm1s_filtered%>%select(adm1_id, gn_name, geonunit))

# snp_in_maus = snp_study_area%>%
#   st_join(., st_buffer(maus_study_area, 1000))%>%
#   filter(primary_commodity.x == primary_commodity.y)
#   
# snp_filtered = snp_study_area%>%
#   filter(!snp_id %in% snp_in_maus$snp_id)%>%
#   select(-snp_id)

st_write(snp_study_area, "Processed data/Mines/snp_study_area.gpkg", delete_dsn = T)
rm(snp_in_maus, snp)
```

Load the processed mining datasets back in and map to confirm everything looks good
```{r}
maus_filtered = st_read("Processed data/Mines/maus_study_area.gpkg")
snp_filtered = st_read("Processed data/Mines/snp_study_area.gpkg")

tm_shape(maus_study_area%>%filter(geonunit == "Argentina"))+
  tm_polygons(col = "red")+
  tm_shape(st_centroid(maus_study_area%>%filter(geonunit == "Argentina")))+
  tm_dots()+
  tm_shape(snp_study_area%>%filter(geonunit == "Argentina"))+
  tm_dots()+
  tm_basemap("Esri.WorldImagery")
```



# Sampling points

So eventually, we want to match mines to 'non-mines'. 

Let's start off this process by separating our study area in mining areas and non-mining areas.

## Mines

First, let's set a sampling location for each mine.

The S&P dataset is already points so we can sample random points from the Maus dataset (one from each mine) and combine the two.

Sampling one random point per Maus mine
```{r}
# Sample one point per polygon
sampled_pts <- lapply(st_geometry(maus_study_area), function(poly) {
  pt <- st_sample(poly, size = 1, type = "random", exact = TRUE)
  if (length(pt) == 0) st_point(c(NA_real_, NA_real_)) else pt[[1]]})

# Create a new sf object (copying attributes if desired)
maus_sampling_pts = st_sf(geometry = st_sfc(sampled_pts, crs = st_crs(maus_study_area)))%>%
  st_join(., maus_study_area)

maus_sampling_pts = st_centroid(maus_study_area)
  #rename(primary_commodity = Primary.Co)%>%
  #select(primary_commodity, adm1_id:geonunit)%>%
  #mutate(mine_id = row_number(), dataset = "maus")

rm(sampled_pts)
```

Map check to make sure everything looks good
```{r}
tm_shape(maus_study_area%>%filter(gn_name == "State of Tasmania"))+
  tm_borders(col = "red")+
  tm_shape(maus_sampling_pts%>%filter(gn_name == "State of Tasmania"))+
  tm_dots()+
  tm_basemap("Esri.WorldImagery")
```

Combining with deduplicated S&P points to generate 1 dataset with all mining sampling points with covariates
```{r}
comb_mining_sampling_pts = rbind(
  maus_sampling_pts%>%
    mutate(dataset = "maus", primary_commodity = NA), 
  snp_study_area%>%
    mutate(dataset = "snp"))%>%
  mutate(mine_id = row_number())

st_write(comb_mining_sampling_pts, "Processed data/Sampling points/comb_mining_sampling_pts.gpkg", delete_dsn = T)
comb_mining_sampling_pts = st_read("Processed data/Sampling points/comb_mining_sampling_pts.gpkg")
```

Last, let's count how many mines are in each adm1, so that we can calibrate the proportion of how many samples to take in non-mining areas. 
```{r}
sample_n = comb_mining_sampling_pts%>%
  group_by(adm1_id)%>%
  reframe(n_mines = n(), n_sample = n_mines*2)

adm1s_filtered = adm1s_filtered%>%
  left_join(., sample_n, by = "adm1_id")%>%
  filter(!is.na(n_mines))

st_write(adm1s_filtered, "Processed data/Study area/adm1s_filtered.gpkg", delete_dsn = T)
adm1s_filtered = st_read("Processed data/Study area/adm1s_filtered.gpkg")
rm(sample_n, maus_sampling_pts)
```

## Non-Mines

We now want to generate the remaining sampling points as a spatially-balanced sample from non-mining areas. 

To do that, we're going to exclude any areas within 10km of mines. 

So first, let's generate a 10km buffer around each mine. 
```{r}
mine_buffer = rbind(snp_study_area%>%select(geom), maus_study_area%>%select(geom))%>%
  st_buffer(., dist = 10000)

st_write(mine_buffer, "Processed data/Mines/mine_buffer_10km.gpkg", delete_dsn = T)
rm(snp_study_area, maus_study_area)
```

Now we can remove these mining areas to generate non-mine areas within our study adm1s. 
```{r}
non_mine_full = rmapshaper::ms_erase(adm1s_filtered, mine_buffer%>%st_make_valid())%>%
  st_make_valid()

rm(mine_buffer)
```


Now, we want to remove any water from the study area so we don't sample there. 
```{r}
adm1s_no_water = rmapshaper::ms_erase(non_mine_full, water_sf)%>%
  st_make_valid()

st_write(adm1s_no_water, "Processed data/Study area/nonmines_no_water.gpkg", delete_dsn = T)
rm(water_sf)
#adm1s_no_water = st_read("Processed data/Study area/nonmines_no_water.gpkg")
```


Finally, we can sample non-mining areas in the adm1s of interest. For a given adm1, we'll sample 2x the # of mines located in that adm1.
```{r}
adm1s_no_water = adm1s_no_water%>%
  mutate(adm1_id = as.factor(adm1_id), id_pro = paste(gn_name, adm1_id, sep = "_"))

strata_n = setNames(adm1s_no_water$n_sample, adm1s_no_water$id_pro)
non_mining_sample = grts(st_transform(adm1s_no_water, crs_sampling), n_base = strata_n, stratum_var = "id_pro")

non_mining_sample_pts = st_transform(non_mining_sample$sites_base, crs)
st_write(non_mining_sample_pts, "Processed data/Sampling points/non_mining_sampling_pts.gpkg", delete_dsn = T)
rm(non_mining_sample, non_mine_full)

non_mining_sample_pts = st_read("Processed data/Sampling points/non_mining_sampling_pts.gpkg")
```

## Combining
Combine the two.

```{r}
comb_mining_sampling_pts%>%data.frame()
non_mining_sample_pts%>%data.frame()

all_sampling_pts = rbind(
  comb_mining_sampling_pts%>%mutate(stratum = "mine"),
  non_mining_sample_pts%>%select(adm1_id, gn_name, geonunit)%>%mutate(stratum = "non-mine", primary_commodity = NA, dataset = NA, mine_id = NA))%>%
  arrange(adm1_id)%>%
  mutate(point_id = row_number())
  
st_write(all_sampling_pts, "Processed data/Sampling points/all_sampling_pts.gpkg", delete_dsn = T)
all_sampling_pts = st_read("Processed data/Sampling points/all_sampling_pts.gpkg")
```

```{r}
all_sampling_pts%>%data.frame()

tm_shape(all_sampling_pts%>%filter(adm1_id == 10))+
  tm_dots(col = "stratum", palette = "Set1")+
  #tm_basemap("Esri.WorldImagery")+
  tm_shape(adm1s_filtered%>%filter(adm1_id == 10))+
  tm_borders(col = "black")
```



# Extract covariates
```{r}
# Continuous variables
test = all_sampling_pts %>%
  st_join(., landmark_study_area_full%>%select(Name:Doc_Status, documented:indigenous), left = T)%>% 
  group_by(point_id)%>%
  slice(1)%>%
  ungroup()%>%
  
  mutate(indig_no_garnett = ifelse(!is.na(Name), 1, 0))%>%
  
  left_join(., adm1s_full%>%select(adm1_id, landmark_community)%>%st_drop_geometry(), by = "adm1_id")%>%
  
  mutate(
    indigenous = ifelse(is.na(indigenous), 0, indigenous),
    treatment_recog = case_when(
      recognised == 1 ~ 1,
      indigenous == 0 & landmark_community == 1 ~ 0),
    treatment_unrecog = case_when(
      recognised == 0 ~ 1,
      indigenous == 0 & landmark_community == 1 ~ 0),
  
    road_density = terra::extract(road_density, vect(.))[,2],
    pop_density = terra::extract(pop_dens, vect(.))[,2],
    dist_water = terra::extract(dist_to_water, vect(.))[,2],
    forest_cover = terra::extract(forest_binary, vect(.))[,2],
    nightlights = terra::extract(nightlights, vect(.))[,2],
    cropland_cover = terra::extract(cropland_binary, vect(.))[,2],
    protected = 0)
    #protected_areas = as.numeric(st_intersects(., protected_areas_full, sparse=F)),
    
    # dist_to_maus_km = as.numeric(st_distance(geometry, maus$geom[st_nearest_feature(geometry, maus)], by_element = TRUE)) / 1000,
    # dist_to_snp_km = as.numeric(st_distance(geometry, snp$geom[st_nearest_feature(geometry, snp)], by_element = TRUE)) / 1000,
    # dist_to_nearest_mine_km = pmin(dist_to_maus_km, dist_to_snp_km),
    # dist_to_usgs_km = as.numeric(st_distance(geometry, usgs_deposits$geom[st_nearest_feature(geometry, usgs_deposits)], by_element = TRUE)) / 1000,
    # dist_to_mine_or_deposit_km = pmin(dist_to_nearest_mine_km, dist_to_usgs_km))
```

Post-processing with ArcGIS
```{r}
st_write(test, "Processed data/Sampling points/sampling_points_w_covariates.shp", delete_dsn = T)
```

```{r}
protected_arcgis = st_read("Processed data/Sampling points/sampling_points_w_covariates.shp")
test = st_read("Processed data/Sampling points/sampling_points_w_covariates.gpkg")

test = test%>%
  select(-protected)%>%
  left_join(., protected_arcgis%>%select(point_d, protctd)%>%rename(point_id = 1, protected = 2)%>%st_drop_geometry(), by = "point_id")%>%
  mutate(
    across(c(protected, dist_water, pop_density, road_density), ~ if_else(is.na(.), 0, .)),
    stratum_mine = if_else(stratum == "mine", 1, 0))

st_write(test, "Processed data/Sampling points/sampling_points_w_covariates.gpkg", delete_dsn = T)
rm(protected_arcgis)
test = st_read("Processed data/Sampling points/sampling_points_w_covariates.gpkg")
```

Robustness check excluding Maus mines from possible matches
```{r}
test = test%>%
  filter(dataset == "snp" | stratum == "non-mine")
```


# Matching

Set up covariate matching, where for every mine, we find a non-mining point within the same ADM1 that are as similar as possible to the mining point in terms of our other covariates. 

We will run a couple versions of this: 1) for all mines, 2) for mines in areas where we have recognised Indigenous territories, and 3) for mines in areas where we have unrecognised Indigenous territories. 

Notes:
- Try different caliber widths for robustness?
- Try exact match by country (rather than ADM1) as robustness check

1) All mines
```{r}
#Mine vs non-mine
match_all_mine = matchit(
  stratum_mine ~ road_density + pop_density + dist_water + forest_cover + cropland_cover + protected, 
  data = test,
  method = "nearest",
  caliper = 0.1,
  exact = ~ adm1_id)

#Indig vs non-Indig
match_all = matchit(
  indigenous ~ road_density + pop_density + dist_water + forest_cover + cropland_cover + protected,
  data = test,
  method = "nearest",
  caliper = 0.1,
  exact = ~ adm1_id)

#Robustness check of excluding garnett
match_all_no_garnett = matchit(
  landmark_no_garnett ~ road_density + pop_density + dist_water + forest_cover + cropland_cover + protected,
  data = test%>%filter(adm1_id %in% adm1s_no_garnett$adm1_id),
  method = "nearest",
  caliper = 0.1,
  exact = ~ adm1_id)

#save(match_all, file = "match_all_indig.Rdata")
load("match_all_indig.Rdata")
```

2) Mines in areas where we have recognised Indigenous territories
```{r}
match_recog_mine = matchit(
  stratum_mine ~ road_density + pop_density + dist_water + forest_cover + cropland_cover + protected,
  data = test%>%filter(landmark_community == 1 & !is.na(treatment_recog)),
  method = "nearest",
  exact = ~ adm1_id,
  caliper = 0.1)

match_recog = matchit(
  treatment_recog ~ road_density + pop_density + dist_water + forest_cover + cropland_cover + protected,
  data = test%>%filter(landmark_community == 1 & !is.na(treatment_recog)),
  method = "nearest",
  exact = ~ adm1_id,
  caliper = 0.1)

#save(match_recog, file = "match_recog_indig.Rdata")
load("match_recog_indig.Rdata")
```

3) Mines in areas where we have unrecognised Indigenous territories
```{r}
match_unrecog_mine = matchit(
  stratum_mine ~ road_density + pop_density + dist_water + forest_cover + cropland_cover + protected,
  data = test%>%filter(landmark_community == 1 & !is.na(treatment_unrecog)),
  method = "nearest",
  exact = ~ adm1_id,
  caliper = 0.1)

match_unrecog = matchit(
  treatment_unrecog ~ road_density + pop_density + dist_water + forest_cover + cropland_cover + protected,
  data = test%>%filter(landmark_community == 1 & !is.na(treatment_unrecog)),
  method = "nearest",
  exact = ~ adm1_id,
  caliper = 0.1)

#save(match_unrecog, file = "match_unrecog_indig.Rdata")
load("match_unrecog_indig.Rdata")
```

4) Mines in any areas with recognition data
```{r}
match_community_mine = matchit(
  stratum_mine ~ road_density + pop_density + dist_water + forest_cover + cropland_cover + protected,
  data = test%>%filter(landmark_community == 1 & !is.na(recognised)),
  method = "nearest",
  exact = ~ adm1_id,
  caliper = 0.1)

match_community = matchit(
  recognised ~ road_density + pop_density + dist_water + forest_cover + cropland_cover ,
  data = test%>%filter(landmark_community == 1 & !is.na(recognised)),
  method = "nearest",
  exact = ~ adm1_id,
  caliper = 0.1)

#save(match_community, file = "match_community_indig.Rdata")
load("match_community.Rdata")
```
Match on mines: 590 matched points from original dataset of 1569
Match on indig: 332 matched points from original dataset of 1569

## Balance checking

Function to generate balance plots
```{r}
plot_balance <- function(match_data, threshold = 0.1) {
  # Generate the love plot
  love <- love.plot(match_data, threshold = threshold, abs = F)
  
  # Extract and filter the data (removing 'adm1' variables)
  love_data <- ggplot_build(love)
  love_data_filtered <- love_data$plot$data %>%
    filter(!grepl("adm1|geonunit", var) & var != "distance")
  
  # Recreate the plot with both positive and negative threshold lines
  ggplot(love_data_filtered, aes(x = stat, y = var, color = Sample)) +
    geom_point() +
    geom_vline(xintercept = threshold, linetype = 'dashed') +
    geom_vline(xintercept = -threshold, linetype = 'dashed') +
    geom_vline(xintercept = 0) +
    theme_classic() +
    labs(
      x = "Standardized Mean Difference",
      y = "Covariates"
    )
}
```

Create balance plots
```{r}
#summary(match_all, standardize = TRUE)
#bal.tab(match_all, stats = c("mean.diffs", "ks.statistics"))

plot_balance(match_all, threshold = 0.1)
plot_balance(match_recog, threshold = 0.1)
plot_balance(match_unrecog, threshold = 0.1)
plot_balance(match_community, threshold = 0.1)
```

## Summary stats

Match for all lands
```{r}
match_all
match_all_df <- match.data(match_all)
cat("\n", "% of mines successfully matched:", round(nrow(match_all_df%>%filter(stratum == "mine"))/nrow(test%>%filter(stratum == "mine"))*100, 1), "\n", "\n")


match_all_mine
match_all_mine_df <- match.data(match_all_mine)
cat("\n", "% of mines successfully matched:", round(nrow(match_all_mine_df%>%filter(stratum == "mine"))/nrow(test%>%filter(stratum == "mine"))*100, 1))
```
Match on indigenous: 2918 matched points from original dataset of 10509, 25% of mines successfully matched
Match on mine: 4404 matched points from original dataset of 10509, 63% of mines successfully matched

Match for recognised lands:
```{r}
match_recog_mine_df <- match.data(match_recog_mine)


match_recog_df <- match.data(match_recog)
nrow(match_recog_df%>%filter(stratum == "mine"))/nrow(test%>%filter(landmark_community == 1 & !is.na(treatment_recog) & stratum == "mine"))

table(match_recog_df$stratum)
```
Match on indig: 21% of mines successfully matched


Match for unrecognised lands
```{r}
match_unrecog_mine_df = match.data(match_unrecog_mine)

match_unrecog_df <- match.data(match_unrecog)
cat("% of mines successfully matched:", round(nrow(match_unrecog_df%>%filter(stratum == "mine"))/nrow(test%>%filter(landmark_community == 1 & !is.na(treatment_unrecog) & stratum == "mine"))*100, 1))

table(match_unrecog_df$stratum)
```
Match on indig: 10% of mines matched

Recognised vs unrecognised
```{r}
match_community_mine_df = match.data(match_community_mine)
cat("Breakdown of recognition status when matching by mine vs non-mine:")
table(match_community_mine_df$recognised)
```

```{r}
match_community_df <- match.data(match_community)
cat("% of mines successfully matched:", round(nrow(match_community_df%>%filter(stratum == "mine"))/nrow(test%>%filter(landmark_community == 1 & !is.na(recognised) & stratum == "mine"))*100, 1),"\n", "\n")

cat("Breakdown of stratum when matching by recognised vs unrecognised:")
table(match_community_df$stratum)
```



Descriptives requested by Marc for mines vs non-mines:

All Indigenous lands
```{r}
table(match_all_mine_df$indigenous)
```

Recognition treatment
```{r}
table(match_recog_mine_df$treatment_recog)
```

Non-recognition treatment
```{r}
table(match_unrecog_mine_df$treatment_unrecog)
```



# Analysis

## Overall
```{r}
match_all_df <- match.data(match_all)

table(match_all_df$indigenous)
table(match_all_df$stratum)

nrow(match_all_df%>%filter(stratum == "mine"))/nrow(test%>%filter(stratum == "mine"))

model_base = glm(
  indigenous ~ stratum_mine + road_density + pop_density +
                forest_cover + cropland_cover + protected + factor(geonunit),
  data = match_all_df,
  family = binomial())

summary(model_base)
coeftest(model_base, vcov = vcovCL(model_base, cluster = ~ geonunit)) # Clustered SEs by country

# Run McNemar's test
mcnemar.test(table(match_all_df$stratum_mine, match_all_df$indigenous))
chisq.test(match_all_df$stratum_mine, match_all_df$indigenous)

summary(feglm(indigenous ~ stratum_mine + road_density + pop_density +
                forest_cover + cropland_cover + protected | geonunit,
              cluster = "geonunit",
              data = match_all_df,
              family = "binomial"))

summary(feglm(stratum_mine ~ indigenous + road_density + pop_density + forest_cover + cropland_cover + protected | geonunit,
              cluster = "geonunit",
              data = match_all_df,
              family = "binomial"))

summary(feglm(stratum_mine ~ landmark_no_garnett + road_density + pop_density + forest_cover + cropland_cover + protected | geonunit,
              cluster = "geonunit",
              data = match_all_df,
              family = "binomial"))
```
Match okay? 1402 of 5160 points in Indigenous territories. 

Overall, mines are less likely to be on Indigenous territories, compared to the counterfactual of 'non-mines'. Robust to clustering and not clustering standard errors. Also to McNemar's test

```{r}
table(match_all_df$recognised)

summary(feglm(recognised ~ stratum_mine + road_density + pop_density +
                forest_cover + cropland_cover + protected | geonunit,
              cluster = "geonunit",
              data = match_all_df,
              family = "binomial"))
summary(feglm(stratum_mine ~ recognised + road_density + pop_density +
                forest_cover + cropland_cover | geonunit,
              cluster = "geonunit",
              data = match_all_df,
              family = "binomial"))

match_community_df <- match.data(match_community)
table(match_community_df$recognised)
table(match_community_df$stratum)
summary(feglm(recognised ~ stratum_mine + road_density + pop_density + orest_cover + cropland_cover + protected | geonunit,
              cluster = "geonunit",
              data = match_community_df,
              family = "binomial"))

summary(feglm(stratum_mine ~ recognised + road_density + pop_density + forest_cover + cropland_cover + protected | geonunit,
              cluster = "geonunit",
              data = match_community_df,
              family = "binomial"))
```
Mines significantly less likely to be on recognised lands than non-recognised lands. 


## Recognised
Recognised territories vs non-Indigenous lands
```{r}
match_recog_df <- match.data(match_recog)

table(match_recog_df$treatment_recog)
table(match_recog_df$stratum)

model_recog = glm(
  treatment_recog ~ stratum_mine + road_density + pop_density + 
                forest_cover + cropland_cover + protected + factor(geonunit),
  data = match_recog_df,
  family = binomial())

summary(model_recog)

# Clustered SEs by ADM1
coeftest(model_recog, vcov = vcovCL(model_recog, cluster = ~ geonunit))

summary(feglm(treatment_recog ~ stratum_mine + road_density + pop_density +
                forest_cover + cropland_cover + protected | geonunit,
  data = match_recog_df,
  cluster = "geonunit",
  family = "binomial"))

summary(feglm(stratum_mine ~ treatment_recog + road_density + pop_density + forest_cover + cropland_cover + protected | geonunit,
  data = match_recog_df,
  cluster = "geonunit",
  family = "binomial"))

# Run McNemar's test
#mcnemar.test(table(match_recog_df$stratum_mine, match_recog_df$treatment_recog))
#chisq.test(match_recog_df$stratum_mine, match_recog_df$treatment_recog)

```
567 of 2576 observations are on recognised territories.

## Unrecognised
Unrecognised lands vs non-Indigenous lands
```{r}
match_unrecog_df <- match.data(match_unrecog)

table(match_unrecog_df$treatment_unrecog)
table(match_unrecog_df$stratum)

model_unrecog = glm(
  treatment_unrecog ~ stratum_mine + road_density + pop_density + 
                forest_cover + cropland_cover + protected,
  data = match_unrecog_df,
  family = binomial())

summary(model_unrecog)

#coeftest(model_unrecog, vcov = vcovCL(model_unrecog, cluster = ~ geonunit)) # Clustered SEs by ADM1

# Run McNemar's test
mcnemar.test(table(match_unrecog_df$treatment_unrecog, match_unrecog_df$stratum_mine))
#chisq.test(match_unrecog_df$stratum_mine, match_unrecog_df$treatment_unrecog)

summary(feglm(
  treatment_unrecog ~ stratum_mine + road_density + pop_density +  
                forest_cover + cropland_cover + protected | geonunit,
  data = match_unrecog_df,
  family = binomial(),
  cluster = ~geonunit))

match_unrecog_df%>%data.frame()

summary(feglm(
  stratum_mine ~ treatment_unrecog + road_density + pop_density + forest_cover + cropland_cover + protected | geonunit,
  data = match_unrecog_df,
  family = binomial(),
  cluster = ~geonunit))
```
161 of 1579 observations on unrecognised territories. 









# Archive

List of countries for analysis
```{r, eval = F}
countries_df = read.csv("Processed data/data_breakdown_by_country_apr30.csv")

countries_full = as.vector(countries_df%>%filter((landmark %in% c("Full Data", "Indicative areas") | in_garnett == "Yes") & (maus_polygons > 0 | snp_points > 0))%>%pull(country))

countries_recog = as.vector(countries_df%>%filter(landmark == "Full Data" & (maus_polygons > 0 | snp_points > 0))%>%pull(country))
```
