
# Set up

Load packages
```{r}
library(sf)          # Spatial data
library(terra)       # Raster processing  
library(spsurvey)    # Spatially balanced sampling
library(MatchIt)     # Matching
library(cobalt)      # Balance checking
library(tidyverse)       # Data wrangling
library(exactextractr) # Zonal stats
library(rnaturalearth) # Country boundaries
library(tmap)        # Making maps
```

Settings and universal variables
```{r}
tmap_mode(mode = "view")

crs = 3857 # WGS 84 / Pseudo-Mercator (projected)

options(scipen=999)
```

# Import data


## Study area

List of countries
```{r}
countries_df = read.csv("Processed data/data_breakdown_by_country_apr30.csv")

countries_full = as.vector(countries_df%>%filter((landmark %in% c("Full Data", "Indicative areas") | in_garnett == "Yes") & maus_polygons > 0)%>%pull(country))

countries_recog = as.vector(countries_df%>%filter(landmark == "Full Data" & maus_polygons > 0)%>%pull(country))
```


Define study area (based on adm1)
```{r}
#NEED TO GET LIST OF ADM1 WHERE THERE ARE MULTIPLE MINES AND WE HAVE RECOGNITION DATA
```

## Covariates

Road density
```{r}
road_density = rast("Raw data/Covariates/Road access/GRIP4_density_total/grip4_total_dens_m_km2.asc")
```

Distance to water
```{r}
dist_to_water = rast("Raw data/Covariates/Distance to water/distance_to_water.tif")
```

Nightlights
```{r}
nightlights = rast("Raw data/Covariates/Nightlights/VNL_npp_2024_global_vcmslcfg_v2_c202502261200.average_masked.dat.tif/VNL_npp_2024_global_vcmslcfg_v2_c202502261200.average_masked.dat.tif")
```

Protected areas
```{r}
protected_areas = st_read("Raw data/Covariates/Protected areas/WDPA_WDOECM_May2025_Public_all/WDPA_WDOECM_May2025_Public_all.gdb", layer = "WDPA_WDOECM_poly_May2025_all")
```

Land cover
```{r}
land_cover = rast("Raw data/Covariates/Land cover/MCD12C1.A2023001.061.2024251212901.hdf")
```


USGS ECM deposits
```{r}
usgs_deposits = st_transform(st_read("Processed data/usgs_deposits_filtered.gpkg"), crs)%>%
  filter(geounit %in% countries_full)
```

Population Density
```{r}
pop_dens = rast("Raw data/Covariates/Population density/gpw_v4_population_density_rev11_2020_30_sec_2020.tif")
```

Administrative division (adm1) & country
```{r}

```


## Independent variable
```{r}
landmark_simplified = st_transform(st_read(dsn = "C:/Users/benji/Desktop/Landmark/landmark_arcgis/landmark_arcgis.gdb", layer = "community_landmark_simplified", options = "METHOD=SKIP"), crs)%>%
  filter(Country %in% countries_full)%>%
  rename(geometry = Shape)%>%
  st_filter(., buffer, .pred = st_intersects)%>%
  mutate(recognition = ifelse(Form_Rec == "Acknowledged by govt" & Doc_Status == "Documented", 1, 0))
```

## Dependent variable
```{r}
snp = st_transform(st_read("Processed data/snp_points_filtered.gpkg"), crs)%>%
  filter(geounit %in% countries_full)

maus = st_transform(st_read("Processed data/maus_polygons_filtered_joined.gpkg"), crs)%>%
  filter(geounit %in% countries_full)
```


# Generate spatially balanced points

Option A: Stratified by mine proximity (using spsurvey)
```{r}
deposit_buffers$stratum = ifelse(
  st_distance(deposit_buffers, mines) <= 10000, 1, 0  # 10km threshold
)

set.seed(42)
samples = grts(
  sf_object = deposit_buffers,
  n_base = c(1000, 1000),  # Adjust based on your needs
  stratum_var = "stratum"
)
sampled_points = st_as_sf(samples$sites_base)
```

Option B: Halton sequence (uniform random)
```{r, eval = F}
sampled_points = st_halton(2000, crs = st_crs(deposit_buffers)) %>%
  st_as_sf() %>%
  st_intersection(deposit_buffers)
```

# Extract covariates

NEED TO CALCULATE DISTANCE TO NEAREST USGS DEPOSIT POINT
```{r}
# Continuous variables
sampled_points = sampled_points %>%
  mutate(
    road_density = extract(road_density, vect(.))[,2],
    pop_density = extract(pop_density, vect(.))[,2],
    dist_water = extract(dist_water, vect(.))[,2],
    near_mine = st_distance(., mines) <= 10000)

# Land cover classification
lc_reclass = classify(landcover, 
  matrix(c(10,1, 20,2, 30,2), ncol=2, byrow=TRUE))

sampled_points$forest_cover = extract(
  lc_reclass == 1, vect(sampled_points), buffer=1000, fun=mean, na.rm=TRUE)[,2]

sampled_points$farmland_cover = extract(
  lc_reclass == 2, vect(sampled_points), buffer=1000, fun=mean, na.rm=TRUE)[,2]

# Protected areas
sampled_points$protected = as.numeric(
  st_intersects(sampled_points, protected_areas, sparse=FALSE))

# Add country info
sampled_points = st_join(sampled_points, countries)
```

# Matching

```{r}
# Define treatment variables
sampled_points = sampled_points %>%
  mutate(
    indigenous = as.numeric(st_intersects(., indigenous_lands, sparse=FALSE)),
    recognized = as.numeric(st_intersects(., recognized_lands, sparse=FALSE))
  )

# Create control pool
control_pool = filter(sampled_points, indigenous == 0)
```

```{r}
match_recog = matchit(
  recognized ~ road_density + pop_density + dist_water +
               forest_cover + farmland_cover + protected,
  data = bind_rows(
    filter(sampled_points, recognized == 1),
    control_pool
  ),
  method = "nearest",
  exact = ~ country,
  caliper = 0.2,
  ratio = 3
)

matched_recog = match.data(match_recog)
```

Balance checking
```{r}
# Recognized lands balance
bal.tab(match_recog, stats = c("mean.diffs", "ks.statistics"))
love.plot(match_recog, threshold = 0.1, abs = TRUE)

# Non-recognized lands balance  
bal.tab(match_nonrecog, stats = c("mean.diffs", "ks.statistics"))
love.plot(match_nonrecog, threshold = 0.1, abs = TRUE)
```

# Analysis
```{r}
# Add outcome variable
matched_recog = matched_recog %>%
  mutate(mine_outcome = as.numeric(near_mine))

matched_nonrecog = matched_nonrecog %>%
  mutate(mine_outcome = as.numeric(near_mine))

# Recognized lands model
model_recog = glm(
  mine_outcome ~ recognized + road_density + pop_density +
                forest_cover + farmland_cover + factor(country),
  data = matched_recog,
  family = binomial()
)

# Non-recognized lands model
model_nonrecog = glm(
  mine_outcome ~ indigenous + road_density + pop_density +
                forest_cover + farmland_cover + factor(country), 
  data = matched_nonrecog,
  family = binomial()
)

# Compare results
summary(model_recog)
summary(model_nonrecog)
```

