
# Import packages
```{r}
library(sf)          # for vector data
library(terra)       # for raster data
library(exactextractr) # for fast extraction
library(ggplot2)     # for visualization
library(dplyr)       # for data wrangling
library(rnaturalearth)
library(tmap)
library(leaflet)
```

```{r}
crs = 32721
rast_crs = paste("EPSG:", crs, sep = "")

tmap_mode("view")
```


# Import data
```{r}
bolivia = st_transform(ne_countries(country = "Bolivia", scale = "medium", returnclass = "sf"), crs)%>%vect()

road_vector = st_transform(st_read("Raw data/Covariates/Road access/OSM Bolivia (test)/gis_osm_roads_free_1.shp"), crs)

road_density_raster = rast("Raw data/Covariates/Road access/GRIP4_density_total/grip4_total_dens_m_km2.asc")%>%
  project(rast_crs)%>%
  crop(bolivia)%>%
  mask(bolivia)

tm_shape(road_density_raster)+
  tm_raster()
```




# Distance to roads

## Based on road density raster

Create binary road mask (1 = road, NA = no road) from road density raster.
Then calculate distance to roads based on that binary road mask
```{r}
road_mask <- classify(road_density_raster, 
                     rcl = matrix(c(0, NA, 
                                   0.0001, Inf, 1), 
                                 ncol = 2, byrow = TRUE), 
                     right = FALSE)
                     
distance_raster_from_density <- distance(road_mask)%>%
  mask(bolivia)
```

## Based on road vectors

Generating distance to roads raster from road vector sf
```{r}
# Create a blank raster with the same properties as road density
template = rast(road_mask)
  
# Rasterize the roads (1 for roads, NA for non-roads)
roads_rasterised <- rasterize(road_vector, template, field=1, background=0)
  
# Calculate distance to nearest road (in meters if CRS is projected)
distance_raster_from_vector = distance(roads_rasterised)%>%
  mask(bolivia)
```

## Map Comparison
```{r}
my_breaks <- c(0, 10000, 20000, 30000, 40000, 50000)

pal <- colorBin(
  palette = c("#2b83ba", "#80bfac", "#d5eea0", "#ffffbf", "#fdc980", "#f88d51", "#d7191c"),
  domain = my_breaks,
  bins = my_breaks,
  na.color = "transparent")

leaflet() %>%
  addTiles() %>%
  addRasterImage(distance_raster_from_vector, color = pal, group = "Distance Derived from Road Vector") %>%
  addRasterImage(distance_raster_from_density, color = pal, group = "Distance Derived from Density") %>%
  addLegend(pal = pal, values = values(distance_raster_from_vector), title = "Distance", position = "bottomright")%>%
  addLayersControl(
    overlayGroups = c("Distance Derived from Road Vector", "Distance Derived from Density"),
    options = layersControlOptions(collapsed = FALSE))

diff_raster <- distance_raster - distance_raster_mask
plot(diff_raster, main = "Difference (m)")
```



```{r}
# Create sample points for validation (10,000 stratified random points)
set.seed(123)

vals <- spatSample(c(distance_raster, distance_raster_mask), 10000, method="random")
vals <- na.omit(vals)

cor(vals[,1], vals[,2])^2
```

# Alternative metrics

Alternative metric idea: take log of road density so we compress high urban density while stretching rural variation
```{r}
connectivity <- log(road_density_raster + 0.001)

tm_shape(connectivity)+
  tm_raster()
```
Similar: 1/road density
```{r}
inverse_density = 1/road_density_raster

tm_shape(inverse_density)+
  tm_raster()
```


